The Leaky Semicolon

### 2021-10-25 paper/abstract.tex
Integrating referee comments.

% Program logics and semantics tell us that in order to derive
% \begin{math}
%   ((S_1\SEMI S_2), \sigma_0) \Downarrow \sigma_2,
% \end{math}
% we derive 
% \begin{math}
%   (S_1, \sigma_0) \Downarrow \sigma_1
% \end{math}
% and then
% \begin{math}
%   (S_2, \sigma_1) \Downarrow \sigma_2.
% \end{math}
Program logics and semantics tell us that when executing $(S_1; S_2)$ starting in state $s_0$,
we execute $S_1$ in $s_0$ to arrive at $s_1$, then execute $S_2$ in $s_1$ to
arrive at the final state $s_2$.  This is, of course, an abstraction.  Processors
execute instructions out of order, due to pipelines and caches, and compilers
reorder programs even more dramatically.  All of this reordering is meant to
be unobservable in single-threaded code, but is observable in multi-threaded code.
A formal attempt to understand the resulting mess is known
as a ``relaxed memory model.''  The relaxed memory models that have been
proposed to date either fail to address sequential composition directly, or overly
restrict processors and compilers.

To support sequential composition while targeting modern hardware, we propose
using preconditions and families of predicate transformers.  When composing
$(S_1;S_2)$, the predicate transformers used to validate the preconditions of
events in $S_2$ are chosen based on the semantic dependencies from events in
$S_1$ to events in $S_2$.  We apply this approach to two existing memory
models: ``Modular Relaxed Dependencies'' for C11 and ``Pomsets with
Preconditions.''

>>>>

% Program logics and semantics tell us that in order to derive
% \begin{math}
%   ((S_1\SEMI S_2), \sigma_0) \Downarrow \sigma_2,
% \end{math}
% we derive 
% \begin{math}
%   (S_1, \sigma_0) \Downarrow \sigma_1
% \end{math}
% and then
% \begin{math}
%   (S_2, \sigma_1) \Downarrow \sigma_2.
% \end{math}
Program logics and semantics tell a pleasant story about sequential
composition: when executing $(S_1; S_2)$, we first execute $S_1$ then $S_2$.
To improve performance, however, processors execute instructions out of
order, and compilers reorder programs even more dramatically.  By design,
single-threaded systems cannot observe these reorderings; however,
multiple-threaded systems can, making the story considerably less pleasant.
A formal attempt to understand the resulting mess is known as a ``relaxed
memory model.''  Prior models either fail to address sequential composition
directly, or overly restrict processors and compilers, or permit nonsense
thin-air behaviors which are unobservable in practice.

To support sequential composition while targeting modern hardware, we enrich
the standard event-based approach with \emph{preconditions} and
\emph{families of predicate transformers}.
%
% When calculating the precondition of an event $e$ in $\sem{S_1;S_2}$, from
% $\sem{S_2}$ is derived by applying a predicate transformer chosen based on
% events of $\sem{S_1}$ upon which $e$ depends.
When calculating $\sem{S_1;S_2}$, the predicate transformer applied to the
precondition of an event $e$ from $\sem{S_2}$ is chosen based on the set of
events in $\sem{S_1}$ upon which $e$ depends.  We apply this approach to two
existing memory models.
% : ``Modular Relaxed Dependencies'' for C11 and
% ``Pomsets with Preconditions.''
